---
title: "Fig3_analysis"
author: "C-M Svensson"
date: '2022-05-16'
output: html_document
---

Copyright 2022 by Dr. Carl-Magnus Svensson

Research Group Applied Systems Biology - Head: Prof. Dr. Marc Thilo
Figge
https://www.leibniz-hki.de/en/applied-systems-biology.html
HKI-Center for Systems Biology of Infection
Leibniz Institute for Natural Product Research and Infection Biology -
Hans Knöll Insitute (HKI)
Adolf-Reichwein-Straße 23, 07745 Jena, Germany

Code is available for redistribution under the BSD3 license 
(https://opensource.org/licenses/BSD-3-Clause).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = 'svg')
library(tidyverse)
library(dplyr)
library(ggplot2)
library(reshape)
library(forcats)
library(RColorBrewer) 
library(Hmisc)
```

## Entropy

Here I will first create Fig 3 from the manuscript and analyse the differences in terms of
activity level and specificity. The overall change in activity will be measured as the entropy of the 
histograms for the overall activity. A low entropy value indicate that the overall activity
changes a lot when the position is changed. In most cases a low entropy means that
activity is very low what ever change from WT that are made. See for example positions
759 and 728. Positions with an entropy above 3 bits, do produce a large amount of substrate
for a number of different residue. 


```{r Load data, echo=FALSE}
activity = read.csv('./Data/Library.csv', sep=',')
substrates = c( "AlaHA","CysHA", "IleHA","LeuHA","DPheHA", "LPheHA", "DValHA", "LValHA","MetHA", "ThrHA","TrpHA", "TyrHA")
residues_of_interest = c("G", "A", "S", "T", "V", "C", "L", "I", "M", "F", "Y", "W", "R", "E", "D", "Q", "K", "P", "N", "H")
activity_cleaned <- activity[,substrates]
activity_cleaned["Residue"] = activity$Residue
activity_cleaned["Source"] = activity$Source
activity_cleaned["Position"] = activity$Position
activity_cleaned["WT"] = activity$WT

activity_cleaned <- activity_cleaned[activity_cleaned$Residue %in% residues_of_interest, ]
# Read the BLOSUM matrix
file <- "ftp://ftp.ncbi.nih.gov/blast/matrices/BLOSUM62"
b62 <- as.matrix(read.table(file, check.names=FALSE))

```

```{r Entropy}
  entropy <- function(histo) {
    norm_histo = histo/sum(histo, na.rm=TRUE)
    e = -sum(norm_histo*log2(norm_histo), na.rm=TRUE)
    return(e)
  }
```

## Kullback-Leibler divergence

To detect specific mutations that change the profile of produced substrates, we 
initially calculated the Kullback-Leibler divergence (KLd) between the histogram 
of WT produced substrates and all the possible mutation at each position. However,
KLd is a non-symmetric measure that can give some strange behavior. For example,
the KLd V->W gave a KLd of ~1.7 while KLd for W->V is ~9.9. Therefore, we changed
to the symmetric Jensen-Shannon distance (JSD). Besides the symmetry, is JSD 
bounded between 0 and 1, giving an intuitive feel for the difference of the 
distributions. To visualize this, we have plotted an normalized version of 
Figure 3 in the manuscript where each total activity is set to 1. This highlights 
the different substrate profiles we get better. However, sometimes mutations with
low activity have a very different substrate distribution that might be because 
of measuring errors. See for example the insertion of residues K or P in position 
654. These give entirely different substrate distributions than the WT S-residue, 
but the overall activity is so low that the bars are not even visible when looking 
at the un-normalized bar plot. Therefore I required that the total substrate 
production should be at least 1% of the WT production to be considered for JSD 
calculation. I then foucs on the highest JSD between WT and a mutation as a 
measure for the profile change when mutating a certain position. The
threshold of 1% was chosen by me after discussion with Hajo and can be changed 
if a different value makes more sense.

```{r Kullback–Leibler divergence}
  KL_div <- function(h1, h2) {
    kld = sum(h1*(log2(h1/h2)), na.rm=TRUE)
    return(kld)
  }
```
```{r Jensen-Shannon distance}
  JS_dist <- function(h1, h_WT) {
    M = 0.5*(h1/sum(h1, na.rm=TRUE) + h_WT/sum(h_WT, na.rm=TRUE))
    #print(M)
    JSD = 0.5*KL_div(h_WT,M) + 0.5*KL_div(h1,M)
    if (JSD > 0){
      JSD = sqrt(JSD)
    }
    return(JSD)
  }

```
## Loop over the binding positions and determine the activity changes for each residue

```{r normalize activity, echo=FALSE}
  # Setting up data and plotting
  positions = unique(activity_cleaned$Position)
  cols = c("Position", "WT", "Entropy", "WT_G", "WT_A", "WT_S", "WT_T",  "WT_V",  "WT_C",  "WT_L",  "WT_I", "WT_M", "WT_F", "WT_Y", "WT_W", "WT_R", "WT_E", "WT_D", "WT_Q", "WT_K", "WT_P", "WT_N", "WT_H", "max_JS", "BLOSUM", "Full_pos") 
  output_df = setNames(data.frame(matrix(ncol = 26, nrow = 1)), cols)
  HA_colors <- c("#f57268", "#00bd7d", "#a1a300ff", 
               "#0047bbff", "#a8b6c0ff", "#ee1187ff", 
               "#ff8ccdff", "#797ecbff", "#00bce2ff",
               "#1a1a1aff", "#8dd3c7ff", "#c059cbff")
  HA<-c("AlaHA","CysHA", "IleHA",
      "LeuHA","DPheHA", "LPheHA", 
      "DValHA", "LValHA","MetHA",
      "ThrHA","TrpHA", "TyrHA")
  CB = c("WT","Residue", "BLOSUM", "JSD", "Position")
  BLOSUM_DF = setNames(data.frame(matrix(ncol = 5, nrow = 1)), CB)

  combined_act_pos_norm = NULL
  combined_act_pos_norm_per_enzyme = NULL
  for (pos in positions){
    act_pos = subset(activity_cleaned, Position==pos)
    
    WT = unique(act_pos$WT)
    out = c(pos, WT)
    act_pos_norm = NULL
    for (s in unique(act_pos$Source)){
      act_pos_s = subset(act_pos, Source==s)
      if (s == "MIS1" | s == "MIS2" | s == "APPEND"){
        act_pos_WT = subset(activity, Source==s)
        act_pos_WT = subset(act_pos_WT, Residue=="WT")
        norm = mean(rowSums(act_pos_WT[,substrates]))
      }
      else {
        act_pos_WT = subset(act_pos_s, Residue==WT)
        norm = mean(rowSums(act_pos_WT[,substrates]))
      }

      if (is.null(act_pos_norm)){
        act_pos_norm = act_pos_s[,substrates]/norm
        act_pos_norm['Residue'] = act_pos_s$Residue
      }
      else{
        act_pos_norm_tmp = act_pos_s[,substrates]/norm
        act_pos_norm_tmp['Residue'] = act_pos_s$Residue
        act_pos_norm = rbind(act_pos_norm, act_pos_norm_tmp)
      }
    }
    act_pos_norm = aggregate(act_pos_norm[, 1:12], list(act_pos_norm$Residue), mean)
    names(act_pos_norm)[names(act_pos_norm) == "Group.1"] <- "Residue"
    act_pos_norm_long = melt(act_pos_norm)
    act_pos_norm["Total_activity"] = rowSums(act_pos_norm[,substrates])
    act_pos_norm["Position"] = pos
    act_pos_norm["WT"] = WT
    if (is.null(combined_act_pos_norm)){
      combined_act_pos_norm = act_pos_norm
    }
    else{
      combined_act_pos_norm = rbind(combined_act_pos_norm, act_pos_norm)
    }
    ent = entropy(act_pos_norm$Total_activity)
    out = c(out, ent)
    title = paste(WT,pos, sep="")
    title = paste(title, ", Entropy=", sep="")
    title = paste(title, toString(ent), sep="")
    act_pos_norm_long$variable <- factor(act_pos_norm_long$variable, levels = HA)
    plot_tit = paste("./Plots/relative_activity_", pos,sep="")
    plot_tit = paste(plot_tit, ".svg",sep="")
    ggp <- ggplot(act_pos_norm_long, aes(x=Residue, y=value, fill = variable))+geom_col() + scale_x_discrete(limits = residues_of_interest) + ggtitle(title) + theme_bw() + labs(x="Residue", y = "Relative activity") +
  scale_fill_manual(values=HA_colors) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file=plot_tit, plot=ggp, width=5, height=3)
    print(ggp)
    
    act_pos_norm_per_enzyme = subset(act_pos_norm, select=-Total_activity)
  
    names(act_pos_norm_per_enzyme)[names(act_pos_norm_per_enzyme) == "Group.1"] <- "Residue"
    for (res in unique(act_pos_norm_per_enzyme$Residue)){
      act_pos_norm_per_enzyme[act_pos_norm_per_enzyme$Residue==res, substrates] = act_pos_norm_per_enzyme[act_pos_norm_per_enzyme$Residue==res, substrates]/sum(act_pos_norm_per_enzyme[act_pos_norm_per_enzyme$Residue==res, substrates])
    }
    
    
    act_pos_norm_per_enzyme_long = melt(act_pos_norm_per_enzyme)
    act_pos_norm_per_enzyme["Total_activity unnormalized"] = rowSums(act_pos_norm[,substrates])
    act_pos_norm_per_enzyme["Position"] = pos
    act_pos_norm_per_enzyme["WT"] = WT
    if (is.null(combined_act_pos_norm_per_enzyme)){
      combined_act_pos_norm_per_enzyme = act_pos_norm_per_enzyme
    }
    else{
      combined_act_pos_norm_per_enzyme = rbind(combined_act_pos_norm_per_enzyme, act_pos_norm_per_enzyme)
    }
    
    title = paste(WT,pos, sep="")
    title = paste(title, ", Max JSD=", sep="")
    WT_hist = act_pos_norm_per_enzyme[act_pos_norm_per_enzyme$Residue==WT, substrates]
    max_JS = 0.0
    RI = ""
    for (r in residues_of_interest){
        TA_r = act_pos_norm[act_pos_norm_per_enzyme$Residue==r, "Total_activity"]
        hist = act_pos_norm_per_enzyme[act_pos_norm_per_enzyme$Residue==r, substrates]
        js = JS_dist(WT_hist,hist)
        out = c(out, js)
        if (!is.na(TA_r) &TA_r>0.01){
          BLOSUM_DF = rbind(BLOSUM_DF, c(WT,r,b62[WT,r], js, pos))
          if (js > max_JS){
            max_JS = js
            RI = r
          }
        }
     }
    out = c(out, max_JS)
    blo = b62[WT,RI]
    out = c(out, blo)
    full_name = paste(WT, pos, sep="")
    full_name = paste(full_name, RI, sep="")
    out = c(out, full_name)
    output_df = rbind(output_df, out)
    title = paste(title, toString(max_JS), sep="")
    title = paste(title, ", for residue ", sep="")
    title = paste(title, RI, sep="")
    act_pos_norm_per_enzyme_long$variable <- factor(act_pos_norm_per_enzyme_long$variable, levels = HA)
    plot_tit = paste("./Plots/activity_distribution", pos,sep="")
    plot_tit = paste(plot_tit, ".svg",sep="")
    ggp2 <- ggplot(act_pos_norm_per_enzyme_long, aes(x=Residue, y=value, fill = variable))+geom_col() + scale_x_discrete(limits = residues_of_interest) + ggtitle(title) + theme_bw() + labs(x="Residue", y = "Normalized total activity") + 
  scale_fill_manual(values=HA_colors) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
    ggsave(file=plot_tit, plot=ggp2, width=5, height=3)
    print(ggp2)
    
  }
  output_df = subset(output_df, !is.na(Position))
  BLOSUM_DF = subset(BLOSUM_DF, !is.na(Position))
  output_df$max_JS = as.numeric(output_df$max_JS)
  output_df$Entropy = as.numeric(output_df$Entropy)
  BLOSUM_DF$BLOSUM = as.numeric(BLOSUM_DF$BLOSUM)
  BLOSUM_DF$JSD = as.numeric(BLOSUM_DF$JSD)
  write.table(combined_act_pos_norm, file="relative_activity.csv", sep = ",", row.names=FALSE)
  write.table(combined_act_pos_norm_per_enzyme, file="normalized_activity.csv",sep = ",", row.names=FALSE)
  
```

```{r WT JSD, echo=FALSE}
  positions = unique(activity_cleaned$Position)
  WT_df_tmp = subset(activity, Position=="WT")
  WT_df <- WT_df_tmp[,substrates]
  WT_df["Residue"] = WT_df_tmp$Residue
  WT_df["Source"] = WT_df_tmp$Source
  WT_df["Position"] = WT_df_tmp$Position
  WT_df["WT"] = WT_df_tmp$WT
  for (pos in positions){
    act_pos = subset(activity_cleaned, Position==pos)
    
    WT_s = unique(act_pos$WT)
    WT_df_tmp = subset(act_pos, Residue==WT_s)
    WT_df = rbind(WT_df, WT_df_tmp)
  }

  for (res in unique(WT_df$Residue)){
      WT_df[WT_df$Residue==res, substrates] = WT_df[WT_df$Residue==res, substrates]/rowSums(WT_df[WT_df$Residue==res, substrates])
  }
  WT_js = NA
  for (row1 in 1:(nrow(WT_df)-1)){
    for (row2 in (row1+1):nrow(WT_df)){
      if (WT_df[row1,"Source"] == WT_df[row2,"Source"]){
      js = JS_dist(WT_df[row1,substrates],WT_df[row2,substrates])
      if (row1 == 1 && row2==2){
        WT_js = js
      }
      else{
         WT_js = c(WT_js,js)
      }
      }
    }
  }
```

## Entropy vs. JSD 
Below is a 2D scatter plot with Entropy and maximal JSD for each position. We
see that the Entropy is pretty uniformly spread between 1 and 4 bits. JSD is also
much more spread compared to the old KLd measure. The line in the plot indicates
the 95th percentile level of JSD between WT experiments. I compared the JSDs 
between all WT experiments that are from the same Source. I did not compare
JSD for WTs from different plates.

```{r Scatter plot, echo=FALSE}
   sec_shell = c("654", "658", "661", "664", "703", "727", "759")
   spec_code = c("660", "663", "702", "726", "728", "752", "760", "761")
   cols = c("Position", "WT", "Entropy", "WT_G", "WT_A", "WT_S", "WT_T",  "WT_V",  "WT_C",  "WT_L",  "WT_I", "WT_M", "WT_F", "WT_Y", "WT_W", "WT_R", "WT_E", "WT_D", "WT_Q", "WT_K", "WT_P", "WT_N", "WT_H", "max_JS", "BLOSUM", "Full_pos", "Shell")

  shell_colors = c("#000000", "#00FFFF", "#808080")
  output_df = output_df %>%  mutate(Shell = NA)
  output_df[output_df$Position %in% sec_shell, "Shell"] = "Second shell"
  output_df[output_df$Position %in% spec_code, "Shell"] = "Specificity code"
  
  WT_js_df =data.frame(WT_js)
  WT_js_df = WT_js_df %>%  mutate(Shell = "WT", Position="WT")
  WT_95q = quantile(WT_js_df$WT_js,0.95)
  WT_95q_df =data.frame(WT_95q)
  WT_95q_df = WT_95q_df %>%  mutate(Shell = 1)
  add = c(WT_95q, 4)
  WT_mean = mean(WT_js_df$WT_js)
  WT_mean_df = data.frame(WT_mean)
  
  WT_95q_df = WT_95q_df %>%  mutate(Position="WT")
  WT_95q_df = rbind(WT_95q_df, add)
  WT_95q_df$WT_95q = as.numeric(WT_95q_df$WT_95q)
  output_df$Entropy = as.numeric(output_df$Entropy)
  output_df$max_JS = as.numeric(output_df$max_JS)
  output_df$BLOSUM = as.numeric(output_df$BLOSUM)
  
  ggp3 = ggplot(output_df, aes(Entropy, max_JS, color=Shell)) + geom_point() + 
  geom_text(label=output_df$Full_pos, nudge_x = 0.1, nudge_y = 0.05, size=3) + ylim(0.0, NA) +
  theme_bw() + labs(x="Entropy", y="Maximal JSD") + geom_line(data=WT_95q_df, aes(x=Shell, y=WT_95q, colour="WT JSD 95% line", alpha=0.4), size=1, linetype="dashed") + scale_colour_manual(values=shell_colors) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 5),
        axis.text.x = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  
  ggsave(file="./Plots/max_JSD_vs_Entropy_small.svg", plot=ggp3, width=2.6, height=1.8)
  print(ggp3)
```

## Difference between second shell and Specificity code
It is clear that the Entropy (i.e. overall activity change) is pretty similar 
between both positions. This is also not a significant difference. The KL divergence 
for the specificity codes are on the other hand higher for a number of positions. 
The difference is currently not significant, but one could maybe include some more
data than just means to see if that improves the situation.

```{r Shell analysis, echo=FALSE}
  WT_95q = quantile(WT_js_df$WT_js,0.95)
  WT_95q_df =data.frame(WT_95q)
  WT_95q_df = WT_95q_df %>%  mutate(Shell = 0.5)
  add = c(WT_95q, 2.5)
  WT_95q_df = WT_95q_df %>%  mutate(Position="WT")
  WT_95q_df = rbind(WT_95q_df, add)
  WT_95q_df$WT_95q = as.numeric(WT_95q_df$WT_95q)
  
  pJS = t.test(max_JS~Shell, data = output_df)
  title_js = "t-test gives significance value "
  title_js = paste(title_js, pJS$p.value)
  ggp4 = ggplot(output_df, aes(Shell, max_JS, colour=Position)) + geom_jitter(position=position_jitter(0.2), size=5) + ggtitle(title_js) + geom_line(data=WT_95q_df, aes(x=Shell, y=WT_95q, colour="WT JSD 95% line"),size=3) +
    theme_bw() + labs(x="Shell", y="Maximal JSD")
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file="./Plots/max_JSD_per_shell.svg", plot=ggp4, width=5, height=3)
  print(ggp4)
  
  pE = t.test(Entropy~Shell, data = output_df)
  titleE = "t-test gives significance value "
  titleE = paste(titleE, pE$p.value, sep="")
  ggp5 = ggplot(output_df, aes(Shell, Entropy, colour=Position)) + geom_jitter(position=position_jitter(0.2), size=5)+ ggtitle(titleE) + theme_bw() + labs(x="Shell", y="Entropy") +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file="./Plots/Entropy_per_shell.svg", plot=ggp5, width=5, height=3)
  print(ggp5)
  knitr::kable(
    output_df,
    col.names = cols
  )
  
```

## Max JSD values per position with BLOSUM values
If we plot the maximal JSD vs. the BLOSUM values for these replacements we see 
that all BLOSUM values are <0. 

```{r Max BLOSUM values vs. JSD, echo=FALSE}
  set.seed(420)

    shell_colors = c("#000000", "#00FFFF", "#808080")
    max_blosum_plot2 = ggplot(output_df, aes(BLOSUM, max_JS, colour=Shell))+ geom_text(label=output_df$Full_pos,position=position_jitter(width=0.4,height=0.025))+ theme_bw() +   labs(x="BLOSUM value", y="JSD") +
  geom_line(data=WT_95q_df, aes(x=c(-3.5,-0.5), y=WT_95q, colour="WT JSD 95% line", alpha=0.4), size=3,linetype="dashed") + ylim(0.0, NA) +   scale_colour_manual(values=shell_colors) +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file="./Plots/max_JSD_vs_BLOSUM.svg", plot=max_blosum_plot2, width=5, height=3)
  print(max_blosum_plot2)
```

## All JSD values (above 1% total activity) and BLOSUM values
We see that if we have a BLOSUM value >0, we will not have JSD value that is 
larger than what we will see between different WT experiments.

```{r All BLOSUM values vs. JSD, echo=FALSE}
  B_DF = subset(BLOSUM_DF, WT != Residue)
  blosum_plot = ggplot(B_DF, aes(BLOSUM, JSD, colour=Position))+geom_jitter(position=position_jitter(0.1),size=3)+ theme_bw() + labs(x="BLOSUM value", y="JSD") +
  geom_line(data=WT_95q_df, aes(x=c(-4,3), y=WT_95q, colour="WT JSD 95% line"), size=3, linetype="dashed") +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        #legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file="./Plots/all_JSD_vs_BLOSUM.svg", plot=blosum_plot, width=5, height=3)
  print(blosum_plot)
  BLOSUM_DF = BLOSUM_DF %>%  mutate(Shell = NA)
  BLOSUM_DF[BLOSUM_DF$Position %in% sec_shell, "Shell"] = "Second shell"
  BLOSUM_DF[BLOSUM_DF$Position %in% spec_code, "Shell"] = "Specificity code"
  B_DF = subset(BLOSUM_DF, WT != Residue)
  shell_colors = c("#000000", "#00FFFF", "#808080")
  blosum_plot2 = ggplot(B_DF, aes(BLOSUM, JSD, colour=Shell))+geom_jitter(position=position_jitter(0.2),size=1)+ theme_bw() + labs(x="BLOSUM value", y="JSD") +
    geom_line(data=WT_95q_df, aes(x=c(-4,3), y=WT_95q, colour="WT JSD 95% line", alpha=0.4), size=1, linetype="dashed") + scale_colour_manual(values=shell_colors) +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "none", 
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        axis.text.x = element_text(color = "black", size = 9, angle = 45, hjust = 1, vjust = 1, face = "plain"),
        axis.text.y = element_text(color = "black", size = 9, angle = 0, hjust = 1, vjust = 0.5, face = "plain")   
        )
  ggsave(file="./Plots/all_JSD_vs_BLOSUM_shell.svg", plot=blosum_plot2, width=2.6, height=1.8)
  print(blosum_plot2)
```

```{r Linear fits, JSD and BLOSUM, echo=FALSE}
  lm_cols = c("Position", "Intercept", "Slope", "P-value")
  lm_df = NULL
  for (pos in unique(B_DF$Position)){
    B_TMP = subset(B_DF, Position==pos)
    model = lm (JSD~BLOSUM, B_TMP)
    summ = summary(model)
    pv = pf(summ$fstatistic[1], summ$fstatistic[2], summ$fstatistic[3], lower.tail=FALSE)
    int = summ$coefficients['(Intercept)','Estimate']
    est = summ$coefficients['BLOSUM','Estimate']
    if (is.null(lm_df)){
      lm_df = data.frame("Position" = pos, "Intercept" = int, "Slope" = est, "P-value" = pv)
    }
    else{
      lm_df = rbind(lm_df, c(pos, int, est, pv))
    }
    
  }
pl = ggplot(B_DF, aes(x = BLOSUM, y = JSD, color=factor(Position)) ) + geom_point() + geom_smooth(method = "lm", se = FALSE)
  print(pl)
  knitr::kable(
    lm_df,
    col.names = lm_cols
  )
```
